Assessing Amazon Turker and automated machine forecasts in the Hybrid Forecasting Competition

### 2019-01-04 slides/index.Rmd
update slides

---
title: Assessing Amazon Turker and automated machine forecasts in the Hybrid Forecasting Competition
author: "Predictive Heuristics </br>Andreas Beger and Michael D. Ward"
date: "`r format(Sys.Date(), '%e %b %Y')`"
output: 
  xaringan::moon_reader:
    css: ["default-fonts", "default", "custom_theme.css"]
    chakra: libs/remark-latest.min.js
    nature:
      beforeInit: "macros.js"
---

@@

class: center

# Preview

@@

.left.footnote[<sup>1</sup> Individual Forecasting Problem...I may not be able to keep myself from using HFC jargon. </br>
<sup>2 </sup> Hybrid Forecasting Competition </br>
<sup>3 </sup> Randomized controlled trial A]

@@

@@

>The HFC program is developing and testing hybrid geopolitical forecasting systems. These systems **integrate human and machine forecasting components** to create maximally accurate, flexible, and scalable forecasting capabilities.

@@

@@

# The Hybrid part:

# Some users could see time series charts and machine forecasts for some IFPs

@@

- An automated system is collection and updating data from several data sources
- An automated system identifies if data for a particular question is available

If so:

@@

@@

@@

- RCT-A, the first trial period, lasted from 7 March to 7 September 2018
- Use subset of forecasts from 2 May to 1 August 2018

N fcasts, n IFPs

@@

It's not clear why though, e.g.:
- they outperformed on questions that did not have data (charts and model)
- they outperformed volunteer forecasters who saw the machine forecasts even when the machine forecasts had good accuracy

@@

# Model overview

@@

Compare average Brier scores of volunteer, turker, and machine forecasts (reference category)

@@

- Negative values $\rightarrow$ human forecasters did better
- Positive values $\rightarrow$ machine forecasts did better

@@

![:scale 80%](img/model-machine-by-data-source.png)

@@

@@

@@

@@

@@

- OECD interest rates
- FAO food price indices

@@

Register to forecast at:

Internal platform:

>>>>

---
title: Assessing Amazon Turker and automated machine forecasts in the Hybrid Forecasting Competition
author: "Predictive Heuristics </br>Andreas Beger and Michael D. Ward"
date: "`r format(Sys.Date(), '%e %B %Y')` </br>Asian POLMETH 2019, Kyoto, Japan"
output: 
  xaringan::moon_reader:
    css: ["default-fonts", "default", "custom_theme.css"]
    chakra: libs/remark-latest.min.js
    nature:
      beforeInit: "macros.js"
---

@@

class: center, middle

@@

.left.footnote[<sup>1</sup> Individual Forecasting Problem</br>
<sup>2 </sup> Hybrid Forecasting Competition </br>
<sup>3 </sup> Randomized controlled trial A]

@@

---
class: middle

@@

.middle[
>The HFC program is developing and testing hybrid geopolitical forecasting systems. These systems **integrate human and machine forecasting components** to create maximally accurate, flexible, and scalable forecasting capabilities.
]

@@

???

@@

# The hybrid part:

# Some users could see time series charts and/or machine forecasts for some IFPs

@@

- An automated system is collecting and updating data from several data sources, and matching them up, if possible, to questions

If data for a particular question is available:

@@

---
class: middle

### How much crude oil will Iraq produce in May 2018?

.center[
![:scale 80%](img/opec-iraq-with-chart.png)
]

- Less than 4,280
- Between 4,280 and 4,384, inclusive
- More than 4,384 but less than 4,473
- Between 4,473 and 4,576, inclusive
- More than 4,576

---
class: middle

### How much crude oil will Iraq produce in May 2018?

.center[
![:scale 80%](img/opec-iraq-with-forecast.png)
]

- Less than 4,280 *(38%)*
- Between 4,280 and 4,384, inclusive *(15%)*
- More than 4,384 but less than 4,473 *(13%)*
- Between 4,473 and 4,576, inclusive *(13%)*
- More than 4,576 *(21%)*

???

The model is ARIMA(0, 1, 1)

@@

???

Examples of parsing that needs to be done:

- how far ahead to forecast, which requires parsing the data and question time periods
- whether the data are count and should be constrained
- partial data handling
- data aggregation for fixed period questions

# Questions that were of primary concern to us

How well are the machine forecasts doing?

Are there particular question groups where performance is good or lacking? Basically, where do we need to focus improvements?

@@

RCT-A, the first trial period, lasted from 7 March to 7 September 2018

- Use subset of forecasts from 2 May (turkers enter) to 1 August 2018 (change in tracking)

~49,000 forecasts

156 IFPs; 46 with machine forecasts

971 unique users

Forecasters:

- Volunteers who joined the platform 
- Amazon Mechanical Turkers
- Machine (basil-ts)

@@

It's not clear why though, e.g.:

- they outperformed on questions that did not have data (charts and model)

- they outperformed volunteer forecasters who saw the machine forecasts even when the machine forecasts had good accuracy

@@

# Relative forecaster performance

@@

Linear model to compare average Brier scores of volunteer, turker, and machine forecasts (reference category)

@@

- Negative values $\rightarrow$ human forecasters did better

- Positive values $\rightarrow$ machine forecasts did better

@@

![](img/model-machine-by-data-source-buildup.png)

???

- explain the two panels
- x-axis left is average machine forecast performance, right is difference from left; both are Brier scores
- y-axis are question groups, number is number of questions
- sorted so that at top are better human, at bottom are better machine

![](img/model-machine-by-data-source.png)

???

- overall easiest, which can't see directly were oil prices, stocks
- overall hardest were ACLED, OECD interest rates, and Boko Haram
- humans generally tended to do better on questions with monthly period (food prices, oil production)
- machines more clearly did better on questions that require data aggregation (ACLED, Nigeria, hacking); earthquakes aggregates are readily available
- ACLED 2op versus 5op: maybe partly 0's, partly truncation issue

@@

Overall hard questions; machine outperformed human forecasts

> How many battle deaths will ACLED record in Central African Republic in August 2018?

![](img/chart-2117.png)

@@

Overall hard questions; machine has middling performance

> What will be the long-term interest rate for Hungary (HUN) in July 2018?

![](img/chart-1793.png)

@@

Relatively easy questions, but machines underperformed a lot

> Will ACLED record any riot/protest events in Gambia in July 2018?

![](img/chart-1136.png)

@@

- ACLED
- Privacyrights hacking
- Nigeria security tracker (Boko Haram)

@@

- OECD interest rates
- FAO food price indices
- exchange rates
- oil production

# Issues we did not address

What about selection issues; are volunteer forecasters able to self-select into questions they will do well on?

Do the chart volunteers do better on questions requiring data aggregation; generally there are some inconsistencies on why/how the chart volunteers did so well.

@@

Register to forecast at https://sage-platform.isi.edu/
